PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2023-04-05 refereeResponse.txt

5.3 Time-stepping Criterion
554: footnote 2 can just go in the main text body.
	Done
584: footnote 3 can also go in the main text body.
	Done
584: Are there any works on the numerical robustness of this more efficient time-stepping scheme. While I'm inclined to believe that the error is only marginal, I think having some sort of test (or a reference to one) would make the results more steadfast.
	I'm fairly certain that we are the first ones to try this. I've added an appendix that compares this timestepping scheme against a fixed timestep case. There appear to be no meaningful differences. TODO: Show mass distribution as differential or cumulative?

5.4 Results
586: Maybe include reference to figure where this is shown
	Figure 1 of Kokbubo 2002 shows this clearly. A reference to the paper has been added in the text.
592: Is this really always true? Counterexample e.g. Emsenhuber et al 2021 and others.
	The NGPPS models (along with the Clement 2020 paper) seem to be a couple of notable exceptions to this, although the embryo growth is not treated self-consistently leading up to the giant impact phase. We have added a mention of the Emsenhuber papers to the text.
Table 2:
What is the rescaled integration time?
	TODO: Not sure where to explain table columns now?

Figure 5:
I think the bottom panel should have a log y-axis
I would make the color of the dots consistent with Fig.2, where the final outcome is orange
There could be numbers on the contours indicating the density of the initial condition.
The meaning of the dashed line in the bottom panel could be indicated inside the figure itself.
I'm confused by the statement "interior to the oligarchic growth region (around 60d)". Where does this definition come from? Is this defined by bodies exceeding the critical feeding size? If so the statement should be rephrased to make the causal connection clear.
	The y axis on the bottom panel of figure 5 is now log-scaled. Point colors in both panels have been changed to orange, to be consistent with figure 2. The dashed line in the bottom panel is now labeled. Yes, we are defining the oligarchic growth region to be the region where \tilde{b}_required <= 2\sqrt{3}. The last sentence in the figure caption has been rephrased to make this clearer.
606: How can a mass be expressed in units of feeding zone size?
	We are solving equation 11 for \tilde{b} and using the initial local surface density, along with the final particle mass to calculate the feeding zone size. We tried to explain this starting on line 611, but it sounds like some further clarification is needed. Some additional text has been added to the paragraph following equation 11 to explain where the value for \tilde{b} comes from.
620: What is the intuition behind tilde b, how exactly is it defined? Since it is a dimensionless quantity, calling it an orbital separation is not intuitive to me. Can tilde b be converted to a physical size? Also, I struggle to interpret the critical tilde be of 2*sqrt(3). Do I understand correctly that a body with tilde b less than 2 sqrt(3) is associated with a feeding zone so small that close encounters are not possible, and thus are not expected to grow further?
632: Why are the bodies expected to follow the tilde b = 2sqrt(3) from a theoretical perspective? Also, I'm unconvinced that the bodies do that at all in the presented simulations - they fall below this value but don't follow it.
640: If anything the opposite seems to be true, no? Maybe have a specific value for the expected eccentricity of protoplanets with and without inelastic damping.
644: If I look at the contour map by eye, the eccentricity distribution doesn't seem all that different in the inner vs outer disk. Maybe the difference can be quantified?
647: This is an interesting point. I'd be interested in seeing a comparison between inner and outer disk accounting for that difference in time scales. Is it possible to include such a comparison maybe with a figure?
Figure 6:
This figure could be improved if each line was colored by the semi-major axis of the radius bin, and if there was a corresponding colorbar. The difference between the 60 day boundary could still be indicated with an additional line.
665: I struggle to see that there is a significant qualitative difference in the behavior in the lines just inside and just outside the boundary of 60 orbits. The first dashed line also seems to converge to a sigma/Sigma \sim 5 %, not much below the first solid line.
668: add reference. Also footnote 5 can probably go in the main text. Can the authors provide an equation that converts tilde b to units of r_h?
682: This statement seems only to be true for fdshallow. The others have a spread in eccentricity that covers two orders of magnitude...
685: It is not clear to me why more inelastic collisions lead to a more sharply peaked eccentricity distribution, and not just more damping like the authors argue for the fdLo simulation.
701: Again, I don't see that the value for tilde b approaches 2sqrt(3) beyond P=60d orbits. The bodies have smaller feeding zones than tilde b = 2sqrt(3)
702: Are they that similar? Maybe this statement can be more quantitative? To me it looks like the values fall below 2sqrt(3) at 40 days for fdHiShallow and at 70 days for fdLo. The interpretation that the boundary is only driven by the local value for alpha seems somewhat speculative. I'd also be interested how randomness affects this, i.e. if the simulation is initialized with a different random seed, by how much does the final distribution change?

### 2023-04-05 main.tex
Update figure 5

A second issue is introduced when two particles on different timesteps
undergo a collision. As in the previous case, momentum is not
completely conserved because the most recent `kick' steps did not
happen simultaneously for these bodies. Early in the simulation, we
find that this problem tends to trigger runaway growth at the timestep
boundaries first. This issue carries itself forward through the embryo
formation phase, and protoplanets tend to form at the boundaries. To
correct this issue, we ignore collisions between bodies on different
timesteps early in the simulation. We find that preventing
multi-timestep collisions until after the maximum mass grows by a
factor of 10 prevents any artifacts from developing at the timestep
boundaries, while also minimizing the number of `skipped' collisions. \textbf{In the case of the fdHi simulation, only about 20 collisions out of an eventual 900,000 are ignored. To verify that this timestepping scheme does not affect protoplanet growth, we tested an annulus of growing planetesimals with both fixed steps and our two-phase variable timestepping scheme. The results of these tests are shown in appendix \ref{sec:rung_egg}.}

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/fulldisk_e_m_b.png}
    \caption{The final state of the fdHi simulation. In the top panel,
      the contours denote the initial period-eccentricity distribution
      of the planetesimals. Point sizes indicate the relative masses
      of bodies at the end of the simulation. In the bottom panel, we show the
      feeding zone width (see equation \ref{eq:iso}) required to produce
      the final masses of the bodies. The dashed line indicates the feeding
      zone size ($\tilde{b} = 2 \sqrt{3}$) expected for bodies on
      circular orbits.
      Interior to the oligarchic
      growth region (around 60 d), the feeding zone sizes exceed this value.\label{fig:fulldisk_e_m}}
\end{center}
\end{figure}

@@

The timescales for embryo formation depend on
the chosen surface density profile, along with the local orbital
timescale. Protoplanets form first at the inner edge of the disk,
where the dynamical timescales are short. Growth proceeds in an
inside-out fashion, with the outermost regions of the disk completing
the protoplanet assembly phase last \textbf{((as an example, see figure 1 of \citet{kokubo02)}. This radial timescale dependence is not typically
accounted for in planet formation simulations\footnote{Instead, the innermost protoplanets 
patiently `wait' for this phase of evolution to complete everywhere in the disk} \textbf{(a notable exception being \citet{emsenhuber21a, emsenhuber21b})}, and appears to be an
important component to forming realistic solar system analogs
\citep{clement20}. As with the narrow annulus simulations, we stop the
integration once the masses of protoplanets in the outermost region of
the disk reach a steady value. In table \ref{tab:sim_properties}, we
summarize the outcomes of the four ``full disk'' cases.

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is
shown in the orbital period-eccentricity plane. The size of the points
indicates the relative mass of the bodies. In the bottom panel, the
final masses of the bodies (in units of feeding zone size) are shown
as a function of orbital period. The feeding zone size $\tilde{b}$ is
calculated according to \citet{kokubo02}
\begin{equation}\label{eq:iso}
	M = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $M$ is taken to be the final mass of each body. Here, we assume that each embryo has reached its isolation mass, and 
that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the 
size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $
\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more 
clearly visible. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ (indicated by the horizontal dashed line) is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}). In typical oligarchic growth simulations \citep{kokubo98}, protoplanets tend 
to space themselves apart by $\tilde{b} = 10$, although it should be noted that they do not consume all of the planetesimals 
within this distance.

>>>>

A second issue is introduced when two particles on different timesteps
undergo a collision. As in the previous case, momentum is not
completely conserved because the most recent `kick' steps did not
happen simultaneously for these bodies. Early in the simulation, we
find that this problem tends to trigger runaway growth at the timestep
boundaries first. This issue carries itself forward through the embryo
formation phase, and protoplanets tend to form at the boundaries. To
correct this issue, we ignore collisions between bodies on different
timesteps early in the simulation. We find that preventing
multi-timestep collisions until after the maximum mass grows by a
factor of 10 prevents any artifacts from developing at the timestep
boundaries, while also minimizing the number of `skipped' collisions. \textbf{In the case of the fdHi simulation, only about 20 collisions out of an eventual 900,000 are ignored. To verify that this timestepping scheme does not affect protoplanet growth, we tested an annulus of growing planetesimals with both fixed steps and our two-phase variable timestepping scheme. The results of these tests are shown in appendix \ref{sec:rung_ecc}.}

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/fulldisk_e_m_b.png}
    \caption{The final state of the fdHi simulation. In the top panel,
      the contours denote the initial period-eccentricity distribution
      of the planetesimals. Point sizes indicate the relative masses
      of bodies at the end of the simulation. In the bottom panel, we show the
      feeding zone width (see equation \ref{eq:iso}) required to produce
      the final masses of the bodies. The dashed line indicates the feeding
      zone size expected for bodies on circular orbits. \textbf{For the shorter period bodies,
      the feeding zone size exceeds this expected value, which indicates that oligarchic growth is
      not operating here. This boundary occurs near roughly 60 d in orbital period.}\label{fig:fulldisk_e_m}}
\end{center}
\end{figure}

@@

The timescales for embryo formation depend on
the chosen surface density profile, along with the local orbital
timescale. Protoplanets form first at the inner edge of the disk,
where the dynamical timescales are short. Growth proceeds in an
inside-out fashion, with the outermost regions of the disk completing
the protoplanet assembly phase last \textbf{(as an example, see figure 1 of \citet{kokubo02)}}. This radial timescale dependence is not typically
accounted for in planet formation simulations\footnote{Instead, the innermost protoplanets 
patiently `wait' for this phase of evolution to complete everywhere in the disk} \textbf{(a notable exception being \citet{emsenhuber21a, emsenhuber21b})}, and appears to be an
important component to forming realistic solar system analogs
\citep{clement20}. As with the narrow annulus simulations, we stop the
integration once the masses of protoplanets in the outermost region of
the disk reach a steady value. In table \ref{tab:sim_properties}, we
summarize the outcomes of the four ``full disk'' cases.

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is
shown in the orbital period-eccentricity plane. The size of the points
indicates the relative mass of the bodies. In the bottom panel, the
final masses of the bodies (in units of feeding zone size) are shown
as a function of orbital period. The feeding zone size $\tilde{b}$ is
calculated according to \citet{kokubo02}
\begin{equation}\label{eq:iso}
	M = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $M$ is taken to be the final mass of each body. \textbf{The y-values in the bottom panel of figure \ref{fig:fulldisk_e_m} are calculated by solving equation \ref{eq:iso} for $\tilde{b}$, and inputting the initial surface density and final particle mass into the expression.} Here, we assume that each embryo has reached its isolation mass, and 
that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the 
size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $
\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more 
clearly visible. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ (indicated by the horizontal dashed line) is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}). In typical oligarchic growth simulations \citep{kokubo98}, protoplanets tend 
to space themselves apart by $\tilde{b} = 10$, although it should be noted that they do not consume all of the planetesimals 
within this distance.
